A boundary-integral framework to simulate \\ the viscous erosion of a porous medium

### 2018-04-06 docs/methods/erosion_methods.tex
round of revisions

\begin{abstract} 
We develop numerical methods to simulate the fluid-mechanical erosion of many bodies in a two-dimensional Stokes flow. The broad aim is to simulate the erosion of a porous medium (e.g.~ground water flow) with grain-scale resolution. Our fluid solver is based on a second-kind boundary integral formulation of the Stokes equations that is discretized with a spectrally-accurate Nystr\"om method and solved with a fast-multipole-accelerated GMRES. To evolve the individual grains under shear-driven erosion, we regularize via curvature-penalization and solve the associated system using the {\thL} formulation. This framework allows numerically stable treatment of stiff terms, thus permitting large time
steps. The overall accuracy of our method is spectral in space and second-order in time. The method is computationally efficient, with the fluid solver requiring $\mathcal{O}(N)$ operations per GMRES iteration, a mesh-independent number of GMRES iterations, and a one-time $\mathcal{O}(N^2)$ computation to compute the shear stress. We benchmark single-body results against analytical predictions. In our multi-body simulations, we find that erosion leads to spontaneous channelization between bodies in near contact. The channelization is associated with a dramatic reduction in the resistance of the porous medium, much more than would be expected from the reduction in grain size alone.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Contributions}
To perform this simulation of Stokesian erosion of a porous media, two
fast and accurate methods are employed---one to solve the fluid
equations and the other to integrate the eroding grains.  The fluid
equations are reformulated in terms of a boundary integral equation
(BIE)~\cite{poz1992}.  Since the boundary conditions are Dirichlet, we
use a double-layer potential formulation so that the resulting BIE is
second-kind.  This second-kind BIE is discretized with the spectrally
accurate trapezoid rule~\cite{tre-wei2014} which results in a
well-conditioned dense linear system for an unknown density function.
This linear system is solved using fast-multipole-accelerated
GMRES~\cite{saa-sch1986}.  Once the density function is computed, the
shear stress is computed with a spectrally accurate
odd-even quadrature rule~\cite{sid-isr1988}.  In addition, the pressure
and vorticity are computed in the fluid bulk (Fig.~\ref{fig:50bodies})
with an upsampled trapezoid rule in the fluid bulk to assist with
visualizing the effect of erosion in porous media.

%
%To create high-fidelity and computationally efficient numerical methods,
%two fast and accurate methods are employed.  First, the fluid velocity
%and the erosion rate (shear stress) are computed with a
%fast-multipole-accelerated Kryolv method applied to a spectrally
%accurate discretization of the governing equations.  In particular,
%GMRES~\cite{saa-sch1986} is applied to a spectrally-accurate Nystr\"om
%discretization of the BIE~\cite{tre-wei2014}, a spectrally accurate
%quadrature rule is used to compute the shear stress~\cite{sid-isr1988},
%and all arclength derivatives are computed with Fourier differentiation.
%In addition to computing the shear stress, we also compute the vorticity
%in the fluid bulk.  The vorticity is particularly insightful since it is
%identical to the shear stress on no-slip boundaries.  Therefore, by
%computing the vorticity, we are able to identify regions where the
%erosion rate will be fast or slow (Fig.~\ref{fig:50bodies}).

@@

For these porous network simulations, it is essential to resolve the nonlocal effects between the different bodies.  BIEs naturally resolve these interactions, and they have been used extensively to simulate viscous fluids in complex geometries, including porous media flow~\cite{dea-qua-bir-jua2018, bar-mar-vee-zha2016}.  Common challenges of a BIE formulation include developing efficient preconditioners for the discretized system and numerical difficulties associated with nearly touching bodies. In this work, we consider geometries requiring $\mathcal{O}(100)$ GMRES iterations, so that preconditioners are not necessary. In future work, we will address mobile bodies for which near contacts may necessitate the preconditoners. Similarly, near-singular schemes will not be necessary due to the fact that the spacing between receding bodies of fixed position always grows in time. Again, the future consideration of mobile bodies may necessitate these near-singular schemes.

Possible preconditioners include
multigrid~\cite{qua-bir2015a} and the inverse fast multipole
method~\cite{qua-cou-dar2018, cou-pou-dar2017}, but the results
presented in this work require no more than $\mathcal{O}(100)$ GMRES
iterations which is too small to make preconditioners for BIEs
effective; preconditioners will be considered in future work when many
more grains are eroding and being transported by the flow.  
To address the nearly-singular integrands, we use the fact that the grains are
eroding and not moving, so their separation increases with time.
Therefore, in this work, we apply an upsampling scheme, but in future
work, we will require a near-singular integration
scheme~\cite{qua-bir2014a, klo-bar-gre-one2013, bar-wu-vee2015,
hel-oja2008a, bea-lai2001}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Formulation}
\label{s:formulation}
We start by defining the main variables used to model erosion.  We
consider flows that are confined by a solid wall $\Gamma$ that encloses
$M$ eroding bodies as diagrammed in Fig.~\ref{fig:schematic}.  The
bodies are denoted as $\gamma_\ell$, $\ell=1,\ldots,M$, and we write
$\gamma = \gamma_1 \cup \cdots \cup \gamma_M$.  On each body, $\nn$ and
$\ss$ denote the unit normal and unit tangent vectors respectively.
Throughout the paper, we adopt the convention that $\ss$ points in the
counterclockwise direction and $\nn$ points out of the fluid domain
(i.e.~into the bodies). Neglecting inertial forces, the dynamics of the
fluid is fully determined by the position of the bodies $\xx_\ell(s,t)
\in \gamma_\ell$, where $s$ is the arclength and $t$ is time.  Given
$\xx_\ell(s,t)$, $\ell=1,\ldots,M$, derived variables include the fluid
velocity $\uu$, the pressure $p$, and the shear stress $\tau$.   On the
bounding wall, we prescribe a velocity $\UU(\xx,t)$.  Then, the
governing equations are
\begin{equation}
\label{eqn:erosionModel}
\begin{split}
  \mu \Delta \uu = \grad p, &\hspace{20pt} \xx \in \Omega, \gap &&\mbox{conservation
of momentum}\\
\grad \cdot \uu = 0, &\hspace{20pt} \xx \in \Omega, \gap
&&\mbox{\em conservation of mass} \\
\uu = 0, &\hspace{20pt} \xx \in \gamma, \gap &&\mbox{\em no slip on the
bodies} \\
\uu = \UU, &\hspace{20pt} \xx \in \Gamma, \gap &&\mbox{\em outer wall
velocity} \\
\Vn = \abs{\tau}, &\hspace{20pt} \xx \in \gamma,
&&\mbox{\em erosion model},
\end{split}
\end{equation}
where $\Vn$ is the normal velocity of the interface and $\mu$ is the
fluid viscosity.   
The constant of proportionality only sets the time scale, so we have set
$\Vn = |\tau|$ as described in equation~\eqref{eqn:erosionModel}.  For
the sake of numerical stability, we modify this law in
Section~\ref{sec:thetaL}.

The outer wall is formed by rounding off the corners
of $[-3,3] \times [-1,1]$, and we impose the Hagen-Poiseuille flow
\begin{align*}
  \UU(\xx) = \umax (1-y^2,0), \quad \xx \in \Gamma,
\end{align*}
where $U$ sets the maximum of the imposed Poiseuille flow. Often, we
will simply set $U=1$, but in some instances we change $U$ dynamically
to enforce that the pressure drop across the computational cell is
constant as bodies erode within it.  The boundary of the fluid domain is
$\bd\Omega = \gamma_1 \cup \cdots \cup \gamma_M \cup \Gamma$.  To
minimize the boundary effects of the inflow and outflow, we only place
bodies in the center third, $[-1,1] \times [-1,1]$, of the fluid domain.

@@

We describe how the incompressible Stokes equations are solved in
$\Omega$ using a boundary integral equation in Section~\ref{sec:bies}.
In Section~\ref{sec:shearStressLP}, we discuss how the shear stress is
computed.  Finally, in Section~\ref{sec:thetaL}, we describe how corners
in the geometry are avoided and perform a change of coordinates to treat
a stiff term.

@@

Since the double-layer potential is not able to represent force- and
torque-free velocity fields,  we complete the double-layer potential
with Stokeslets and rotlets~\cite{pow-mir1987, pow1993}
\begin{align}
  S[\llambda_\ell,\cc_\ell] = \frac{1}{4\pi} \left(-\log \rho \II + 
    \frac{\rr \otimes \rr}{\rho^2} \right)\llambda_\ell
  \quad \text{and} \quad
  R[\xi_\ell,\cc_\ell] = \frac{\rr^\perp}{\rho^2}\xi_\ell, 
  \quad \ell=1,\ldots,M,
  \label{eqn:stokeslet_rotlet}
\end{align}
where  $\rr = \xx - \cc_\ell$, $\cc_\ell$ is a point inside the
$\ell^{th}$ grainy, $\rr^\perp = (r_2,-r_1)$, and $\rho = \|\rr\|$.
Then, the solution of~\eqref{eqn:stokes} is
\begin{align}
  \label{eqn:completed_DLP}
  \uu(\xx) = \DD[\eeta](\xx) + 
    \sum_{\ell=1}^{M} S[\llambda_\ell,\cc_\ell](\xx) +
    \sum_{\ell=1}^{M} R[\xi_\ell,\cc_\ell](\xx), \quad \xx \in \Omega,
\end{align}
where the density function, Stokeslets, and rotlets satisfy
\begin{subequations}
\label{eqn:completed_BIE}
\begin{align}
  \ff(\xx) &= -\frac{1}{2}\eeta(\xx) + \DD[\eeta](\xx) + 
      \sum_{\ell=1}^{M} S[\llambda_\ell,\cc_\ell](\xx) +
      \sum_{\ell=1}^{M} R[\xi_\ell,\cc_\ell](\xx) + \NN_0[\eeta](\xx),
      \quad \xx \in \bd\Omega, 
      \label{eqn:DLP}\\
  \llambda_\ell &= \frac{1}{2\pi}\int_{\gamma_\ell} \eeta(\yy) \, ds_\yy,
  \quad \ell=1,\ldots,M,
  \label{eqn:stokeslet} \\
  \xi_\ell &= \frac{1}{2\pi}\int_{\gamma_\ell} \yy^\perp \cdot \eeta(\yy)
  \, ds_\yy, \quad \ell=1,\ldots,M.
  \label{eqn:rotlet}
\end{align}
\end{subequations}
Here,
\begin{align*}
  \NN_0[\eeta](\xx) = \int_{\Gamma} 
    (\nn(\xx) \otimes \nn(\yy))\eeta(\yy) \, ds_\yy
\end{align*}
removes the rank one null space resulting from the flux-free requirement
of $\ff$.

This BIE formulation requires that $\bd\Omega$ is smooth.
Unfortunately, smoothness is not guaranteed erosion since corners
develop at locations where the shear stress vanishes.  One option is to
modify our BIE to handle corners~\cite{rac-ser2017,
ser-rok2016, hel2011, gil-hao-mar2014, bre2012}, but these methods
require the corner's location, or a non-uniform spaced mesh, or are
ill-conditioned.  Instead, make a slight modification to the erosion
model in Section~\ref{sec:thetaL} which allows us to control the
smoothness of the grains.  In this manner, the BIE as described above
can be applied.

@@

The deformation tensor of the double-layer
potential~\eqref{eqn:velocityDLP} at an interior point $\xx \in \Omega$
is~\cite{qua-bir2014a}
\begin{equation}
  \label{eqn:shearStressDLP}
  \begin{aligned}
  \ssigma^\DD(\xx) = \frac{1}{2\pi}\int_{\bd\Omega} &\left(
    2\frac{\rr \cdot \nn}{\rho^2} \frac{\rr \cdot \eeta}{\rho^2} \II + 
    \frac{\rr \cdot \eeta}{\rho^4} (\nn \otimes \rr + \rr \otimes \nn) 
    \right. \\
    &\left.
    +\frac{\rr \cdot \nn}{\rho^4} (\eeta \otimes \rr + \rr \otimes \eeta) - 
    8\frac{(\rr \cdot \nn)(\rr \cdot \eeta)}{\rho^6}(\rr \otimes \rr)
  \right) \eeta(\yy) ds_\yy.
  \end{aligned}
\end{equation}
We require the deformation tensor on $\bd\Omega$, and this requires
taking the limit of equation~\eqref{eqn:shearStressDLP} when $\xx$ tends
to $\xx_0 \in \bd\Omega$.  The limiting value is
\begin{align*}
  \lim_{\substack{\xx \rightarrow \xx_0 \\ \xx \in \Omega}}\sigma_\DD(\xx) =
  J[\eeta](\xx_0) + \ssigma^\DD(\xx_0), \quad \xx_0 \in \bd\Omega,
\end{align*} 
where
\begin{align}
  J[\eeta](\xx_0) = \frac{1}{2}\left(\pderiv{\eeta}{\ss} 
    \cdot \ss\right) \left[ 
  \begin{array}{cc}
    s_x^2 - s_y^2 & 2s_x s_y \\ 2s_x s_y & s_y^2 - s_x^2
  \end{array}\right].
  \label{eqn:shearStressJump}
\end{align}
Next, the deformation tensor of the Stokeslets and
rotlets~\eqref{eqn:stokeslet_rotlet} are
\begin{equation}
  \label{eqn:shearStressSR}
  \begin{aligned}
  \ssigma^S(\xx_0) &= \sum_{\ell=1}^{M}
    \frac{\rr \cdot \llambda_\ell}{4\pi\rho^2} \left(
    \II - \frac{2}{\rho^2} \rr \otimes \rr \right),  \\
  \ssigma^R(\xx_0) &= -\sum_{\ell=1}^M
    \frac{\xi_\ell}{\rho^4} \left(\rr \otimes \rr^\perp + 
    \rr^\perp \otimes \rr \right).
  \end{aligned}
\end{equation}
Combining these deformation tensors, the shear stress at $\xx_0 \in
\bd\Omega$ is
\begin{align}
  \tau = -2 \left(J[\eeta](\xx_0) + \ssigma^\DD(\xx_0) + 
    \ssigma^S(\xx_0) + \ssigma^R(\xx_0)\right) \nn \cdot \ss.
  \label{eqn:shearStressLP}
\end{align}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical methods}
\label{s:method}
To perform numerical simulations, we exploit the difference
between the erosion and fluid time scales.  Since erosion occurs
much slower, we freeze the geometry at each time
step, solve the incompressible Stokes equations, and then update the
geometry according to the erosion model.  Therefore, there are three
main numerical methods that need to be developed.  In
Section~\ref{sec:BIE}, we describe how the BIE~\eqref{eqn:completed_BIE}
is solved.  Once the density function is found,
Section~\ref{sec:shearStress} describes how the shear
stress~\eqref{eqn:shearStressLP} is computed.  Then, in
Section~\ref{sec:timeStepping}, a time stepping method for evolving the
interface according to equations~\eqref{eqn:Levo1}
and~\eqref{eqn:thetaevo1} is described.

@@

To numerically compute the completed double-layer
potential~\eqref{eqn:completed_BIE}, we discretize each of the $M$
bodies with  $N_\iin$ points, and the outer boundary with $N_\out$
points.  Then, $N = MN_\iin + N_\out$ is the total number of
discretization points of $\bd\Omega$.  The double-layer potential
in~\eqref{eqn:DLP} is discretized with the trapezoid rule as
\begin{align}
  \DD[\eeta](\xx_i) = \sum_{j=1}^{N} K(\xx_i,\xx_j) \eeta(\xx_j) 
      \Delta s_{j}, \quad i=1,\ldots,N,
  \label{eqn:trapDLP}
\end{align}
where $\Delta s_j$ is the arclength term of $\bd\Omega$ at
$\xx_j$, and
\begin{align*}
  K(\xx,\yy) = \frac{1}{\pi} \frac{\rr \cdot \nn}{\rho^2} 
      \frac{\rr \otimes \rr}{\rho^2},
\end{align*}
where $\rr = \xx - \yy$ and $\rho = \|\rr\|$, is the kernel of the
double-layer potential.  The diagonal term of~\eqref{eqn:trapDLP} is
replaced with the limiting value of the double-layer potential
\begin{align*}
  D(\xx,\xx) = -\frac{\kappa(\xx)}{2\pi}(\ss(\xx) \otimes \ss(\xx)),
\end{align*}
where $\kappa(\xx)$ is the curvature and $\ss(\xx)$ is the unit tangent
vector at $\xx \in \bd\Omega$.  The equations for the
Stokeslets~\eqref{eqn:stokeslet} and the rotlets~\eqref{eqn:rotlet} are
also discretized with the trapezoid rule.  Applying this discretization
strategy to~\eqref{eqn:completed_BIE}, the dense linear system that
needs to be solved is
\begin{equation}
  \label{eqn:completed_BIE_discretized}
  \begin{aligned}
    \ff(\xx_i) &= -\frac{1}{2}\ssigma(\xx_i) + \sum_{j=1}^{N} 
      K(\xx_i,\xx_j) \eeta(\xx_j) \Delta s_{j} + 
      \sum_{\ell=1}^M S[\llambda_\ell,\cc_\ell](\xx_i) \\
      &\hspace{53pt}
      +\sum_{\ell=1}^M R[\xi_\ell,\cc_\ell](\xx_i) 
      +\sum_{\xx_j \in \Gamma} (\nn(\xx_i) \otimes
      \nn(\xx_j))\eeta(\xx_j), \\
    \llambda_\ell &= \sum_{\xx_j \in \gamma_\ell} \eeta(\xx_j) 
      \Delta s_j, \\ 
    \mu_\ell &= \sum_{\xx_j \in \gamma_\ell}
      {(\xx_j - \cc_\ell)}^\perp \cdot \eeta(\xx_j) \Delta s_j,
  \end{aligned}
\end{equation}
for $i=1,\ldots,N$, $\ell=1,\ldots,M$.  We have applied the trapezoid
rule since all the integrands are smooth and periodic, and this
guarantees that the numerical solutions converges to the true solution
with spectral accuracy.

>>>>

\begin{abstract} 
We develop numerical methods to simulate the fluid-mechanical erosion of many bodies in a two-dimensional Stokes flow. The broad aim is to simulate the erosion of a porous medium (e.g.~ground water flow) with grain-scale resolution. Our fluid solver is based on a second-kind boundary integral formulation of the Stokes equations that is discretized with a spectrally-accurate Nystr\"om method and solved with fast-multipole-accelerated GMRES. To evolve the individual grains under shear-driven erosion, we regularize via curvature-penalization and solve the associated system using the {\thL} formulation. This framework allows numerically stable treatment of stiff terms, thus permitting large time
steps. The overall accuracy of our method is spectral in space and second-order in time. The method is computationally efficient, with the fluid solver requiring $\mathcal{O}(N)$ operations per GMRES iteration, a mesh-independent number of GMRES iterations, and a one-time $\mathcal{O}(N^2)$ computation to compute the shear stress. We benchmark single-body results against analytical predictions. In our multi-body simulations, we find that erosion leads to spontaneous channelization between bodies in near contact. The channelization is associated with a dramatic reduction in the resistance of the porous medium, much more than would be expected from the reduction in grain size alone.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Contributions}
To perform this simulation of Stokesian erosion of a porous media, two
fast and accurate methods are employed---one to solve the fluid
equations and the other to integrate the eroding grains.  The fluid
equations are reformulated in terms of a boundary integral equation
(BIE)~\cite{poz1992}.  Since the boundary conditions are Dirichlet, we
use a double-layer potential formulation so that the resulting BIE is
second-kind.  This second-kind BIE is discretized with the spectrally
accurate trapezoid rule~\cite{tre-wei2014} which results in a
well-conditioned dense linear system for an unknown density function,
and the linear system is solved using fast-multipole-accelerated
GMRES~\cite{saa-sch1986}.  Once the density function is computed, the
shear stress is computed with a spectrally accurate
odd-even quadrature rule~\cite{sid-isr1988}.  In addition, the pressure
and vorticity are computed in the fluid bulk (Fig.~\ref{fig:50bodies})
with an upsampled trapezoid rule to assist with visualizing the effect of erosion in porous media.

@@

For these porous network simulations, it is essential to resolve the
nonlocal effects between the different bodies.  BIEs naturally resolve
these interactions, and they have been used extensively to simulate
viscous fluids in complex geometries, including porous media
flow~\cite{dea-qua-bir-jua2018, bar-mar-vee-zha2016}.  Common challenges
of a BIE formulation include developing efficient preconditioners for
the discretized system and numerical difficulties associated with nearly
touching bodies. In this work, we consider geometries requiring
$\mathcal{O}(100)$ GMRES iterations, so that preconditioners are not
necessary. In future work, we will address mobile bodies for which near
contacts may necessitate preconditoners~\cite{qua-bir2015a,
qua-cou-dar2018, cou-pou-dar2017}. Similarly, near-singular schemes will
not be necessary due to the fact that the spacing between receding
bodies of fixed position always grows in time. Again, the future
consideration of mobile bodies may necessitate near-singular
schemes~\cite{qua-bir2014a, klo-bar-gre-one2013, bar-wu-vee2015,
hel-oja2008a, bea-lai2001}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Formulation}
\label{s:formulation}
We start by defining the main variables used to model erosion.  We
consider flows that are confined by a solid wall $\Gamma$ that encloses
$M$ eroding bodies as diagrammed in Fig.~\ref{fig:schematic}.  The
bodies are denoted as $\gamma_\ell$, $\ell=1,\ldots,M$, and we write
$\gamma = \gamma_1 \cup \cdots \cup \gamma_M$.  On each body, $\nn$ and
$\ss$ denote the unit normal and unit tangent vectors respectively.
Throughout the paper, we adopt the convention that $\ss$ points in the
counterclockwise direction and $\nn$ points out of the fluid domain
(i.e.~into the bodies). Neglecting inertial forces, the dynamics of the
fluid is fully determined by the position of the bodies $\xx_\ell(s,t)
\in \gamma_\ell$, where $s$ is the arclength and $t$ is time.  Given
$\xx_\ell(s,t)$, $\ell=1,\ldots,M$, derived variables include the fluid
velocity $\uu$, the pressure $p$, and the shear stress $\tau$.   On the
bounding wall, we prescribe a velocity $\UU(\xx,t)$.  Then, the
governing equations are
\begin{equation}
\label{eqn:erosionModel}
\begin{split}
  \mu \Delta \uu = \grad p, &\hspace{20pt} \xx \in \Omega, \gap &&\mbox{conservation
of momentum}\\
\grad \cdot \uu = 0, &\hspace{20pt} \xx \in \Omega, \gap
&&\mbox{\em conservation of mass} \\
\uu = 0, &\hspace{20pt} \xx \in \gamma, \gap &&\mbox{\em no slip on the
bodies} \\
\uu = \UU, &\hspace{20pt} \xx \in \Gamma, \gap &&\mbox{\em outer wall
velocity} \\
\Vn = \abs{\tau}, &\hspace{20pt} \xx \in \gamma,
&&\mbox{\em erosion model},
\end{split}
\end{equation}
where $\Vn$ is the normal velocity of the interface and $\mu$ is the
fluid viscosity.   
The constant of proportionality of the erosion law only sets the time scale, so we have set
$\Vn = |\tau|$ as described in equation~\eqref{eqn:erosionModel}.  For
the sake of numerical stability, we modify this law in
Section~\ref{sec:thetaL}.

The outer wall is formed by rounding off the corners
of $[-3,3] \times [-1,1]$, and we impose the Hagen-Poiseuille flow
\begin{align}
  \UU(\xx) = \umax (1-y^2,0), \quad \xx \in \Gamma,
\end{align}
where $U$ sets the maximum of the imposed Poiseuille flow. Often, we
will simply set $U=1$, but in some instances we change $U$ dynamically
to enforce that the pressure drop across the computational cell is
constant as bodies erode within it.  The boundary of the fluid domain is
$\bd\Omega = \gamma_1 \cup \cdots \cup \gamma_M \cup \Gamma$.  To
minimize the boundary effects of the inflow and outflow, we only place
bodies in the center third, $[-1,1] \times [-1,1]$, of the fluid domain.

@@

We describe how the incompressible Stokes equations are solved in
$\Omega$ using a boundary integral equation in Section~\ref{sec:bies}.
In Section~\ref{sec:shearStressLP}, we discuss how the shear stress is
computed.  Finally, in Section~\ref{sec:thetaL}, we describe how corners
in the bodies are avoided and perform a change of coordinates to treat
a stiff term.

@@

Since the double-layer potential is not able to represent force- and
torque-free velocity fields,  we complete the double-layer potential
with Stokeslets and rotlets~\cite{pow-mir1987, pow1993}
\begin{align}
  S[\llambda_\ell,\cc_\ell] = \frac{1}{4\pi} \left(-\log \rho \II + 
    \frac{\rr \otimes \rr}{\rho^2} \right)\llambda_\ell
  \quad \text{and} \quad
  R[\xi_\ell,\cc_\ell] = \frac{\rr^\perp}{\rho^2}\xi_\ell, 
  \quad \ell=1,\ldots,M,
  \label{eqn:stokeslet_rotlet}
\end{align}
where  $\rr = \xx - \cc_\ell$, $\cc_\ell$ is a point inside the
$\ell^{th}$ grain, $\rr^\perp = (r_2,-r_1)$, and $\rho = \|\rr\|$.
Then, the solution of~\eqref{eqn:stokes} is
\begin{align}
  \label{eqn:completed_DLP}
  \uu(\xx) = \DD[\eeta](\xx) + 
    \sum_{\ell=1}^{M} S[\llambda_\ell,\cc_\ell](\xx) +
    \sum_{\ell=1}^{M} R[\xi_\ell,\cc_\ell](\xx), \quad \xx \in \Omega,
\end{align}
where the density function, Stokeslets, and rotlets satisfy
\begin{subequations}
\label{eqn:completed_BIE}
\begin{align}
  \ff(\xx) &= -\frac{1}{2}\eeta(\xx) + \DD[\eeta](\xx) + 
      \sum_{\ell=1}^{M} S[\llambda_\ell,\cc_\ell](\xx) +
      \sum_{\ell=1}^{M} R[\xi_\ell,\cc_\ell](\xx) + \NN_0[\eeta](\xx),
      \quad \xx \in \bd\Omega, 
      \label{eqn:DLP}\\
  \llambda_\ell &= \frac{1}{2\pi}\int_{\gamma_\ell} \eeta(\yy) \, ds_\yy,
  \quad \ell=1,\ldots,M,
  \label{eqn:stokeslet} \\
  \xi_\ell &= \frac{1}{2\pi}\int_{\gamma_\ell} \yy^\perp \cdot \eeta(\yy)
  \, ds_\yy, \quad \ell=1,\ldots,M.
  \label{eqn:rotlet}
\end{align}
\end{subequations}
Here,
\begin{align*}
  \NN_0[\eeta](\xx) = \int_{\Gamma} 
    (\nn(\xx) \otimes \nn(\yy))\eeta(\yy) \, ds_\yy
\end{align*}
removes the rank one null space resulting from the flux-free requirement
of $\ff$.

This BIE formulation requires that $\bd\Omega$ is smooth.
Unfortunately, smoothness is not guaranteed by erosion since corners
develop at locations where the shear stress vanishes.  One option is to
modify our BIE to handle corners~\cite{rac-ser2017,
ser-rok2016, hel2011, gil-hao-mar2014, bre2012}, but these methods
require the corner's location, a non-uniform spaced mesh, or are
ill-conditioned.  Instead, make a slight modification to the erosion
model in Section~\ref{sec:thetaL} which allows us to control the
smoothness of the grains.  In this manner, the BIE as described above
can be applied.

@@

The deformation tensor of the double-layer
potential~\eqref{eqn:velocityDLP} at an interior point $\xx \in \Omega$
is~\cite{qua-bir2014a}
\begin{equation}
  \label{eqn:shearStressDLP}
  \begin{aligned}
  \ssigma^\DD(\xx) = \frac{1}{2\pi}\int_{\bd\Omega} &\left(
    2\frac{\rr \cdot \nn}{\rho^2} \frac{\rr \cdot \eeta}{\rho^2} \II + 
    \frac{\rr \cdot \eeta}{\rho^4} (\nn \otimes \rr + \rr \otimes \nn) 
    \right. \\
    &\left.
    +\frac{\rr \cdot \nn}{\rho^4} (\eeta \otimes \rr + \rr \otimes \eeta) - 
    8\frac{(\rr \cdot \nn)(\rr \cdot \eeta)}{\rho^6}(\rr \otimes \rr)
  \right) \eeta(\yy) ds_\yy.
  \end{aligned}
\end{equation}
We require the deformation tensor on $\bd\Omega$, and this requires
taking the limit of equation~\eqref{eqn:shearStressDLP} when $\xx$ tends
to $\xx_0 \in \bd\Omega$.  The limiting value is
\begin{align}
  \lim_{\substack{\xx \rightarrow \xx_0 \\ \xx \in \Omega}}\sigma_\DD(\xx) =
  J[\eeta](\xx_0) + \ssigma^\DD(\xx_0), \quad \xx_0 \in \bd\Omega,
\end{align} 
where
\begin{align}
  J[\eeta](\xx_0) = \frac{1}{2}\left(\pderiv{\eeta}{\ss} 
    \cdot \ss\right) \left[ 
  \begin{array}{cc}
    s_x^2 - s_y^2 & 2s_x s_y \\ 2s_x s_y & s_y^2 - s_x^2
  \end{array}\right].
  \label{eqn:shearStressJump}
\end{align}
Next, the deformation tensor of the Stokeslets and
rotlets~\eqref{eqn:stokeslet_rotlet} are
\begin{equation}
  \label{eqn:shearStressSR}
  \begin{aligned}
  \ssigma^S(\xx_0) &= \sum_{\ell=1}^{M}
    \frac{\rr \cdot \llambda_\ell}{4\pi\rho^2} \left(
    \II - \frac{2}{\rho^2} \rr \otimes \rr \right),  \\
  \ssigma^R(\xx_0) &= -\sum_{\ell=1}^M
    \frac{\xi_\ell}{\rho^4} \left(\rr \otimes \rr^\perp + 
    \rr^\perp \otimes \rr \right).
  \end{aligned}
\end{equation}
Combining these deformation tensors, the shear stress at $\xx_0 \in
\bd\Omega$ is
\begin{align}
  \tau = -2 \left(J[\eeta](\xx_0) + \ssigma^\DD(\xx_0) + 
    \ssigma^S(\xx_0) + \ssigma^R(\xx_0)\right) \nn \cdot \ss.
  \label{eqn:shearStressLP}
\end{align}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical methods}
\label{s:method}
To perform numerical simulations, we exploit the difference
between the erosion and fluid time scales.  Since erosion occurs
much slower, we freeze the geometry at each time
step, solve the incompressible Stokes equations, and then update the
geometry according to the erosion model.  Therefore, there are three
main numerical methods that need to be developed.  In
Section~\ref{sec:BIE}, we describe how the BIE~\eqref{eqn:completed_BIE}
is solved.  Once the density function is found,
Section~\ref{sec:shearStress} describes how the shear
stress~\eqref{eqn:shearStressLP} is computed.  Then, in
Section~\ref{sec:timeStepping}, a time stepping method for evolving the
interface according to equations~\eqref{eqn:Levo1}--\eqref{eq:xsurf}
are described.

@@

To numerically compute the completed double-layer
potential~\eqref{eqn:completed_BIE}, we discretize each of the $M$
bodies with  $N_\iin$ points, and the outer boundary with $N_\out$
points.  Then, $N = MN_\iin + N_\out$ is the total number of
discretization points of $\bd\Omega$.  The double-layer potential
in~\eqref{eqn:DLP} is discretized with the trapezoid rule as
\begin{align}
  \DD[\eeta](\xx_i) = \sum_{j=1}^{N} K(\xx_i,\xx_j) \eeta(\xx_j) 
      \Delta s_{j}, \quad i=1,\ldots,N,
  \label{eqn:trapDLP}
\end{align}
where $\Delta s_j$ is the arclength term of $\bd\Omega$ at
$\xx_j$, and
\begin{align}
  K(\xx,\yy) = \frac{1}{\pi} \frac{\rr \cdot \nn}{\rho^2} 
      \frac{\rr \otimes \rr}{\rho^2},
\end{align}
where $\rr = \xx - \yy$ and $\rho = \|\rr\|$, is the kernel of the
double-layer potential.  The diagonal term of~\eqref{eqn:trapDLP} is
replaced with the limiting value of the double-layer potential
\begin{align}
  D(\xx,\xx) = -\frac{\kappa(\xx)}{2\pi}(\ss(\xx) \otimes \ss(\xx)),
\end{align}
where $\kappa(\xx)$ is the curvature and $\ss(\xx)$ is the unit tangent
vector at $\xx \in \bd\Omega$.  The equations for the
Stokeslets~\eqref{eqn:stokeslet} and the rotlets~\eqref{eqn:rotlet} are
also discretized with the trapezoid rule.  Applying this discretization
strategy to~\eqref{eqn:completed_BIE}, the dense linear system is solved is
\begin{equation}
  \label{eqn:completed_BIE_discretized}
  \begin{aligned}
    \ff(\xx_i) &= -\frac{1}{2}\ssigma(\xx_i) + \sum_{j=1}^{N} 
      K(\xx_i,\xx_j) \eeta(\xx_j) \Delta s_{j} + 
      \sum_{\ell=1}^M S[\llambda_\ell,\cc_\ell](\xx_i) \\
      &\hspace{53pt}
      +\sum_{\ell=1}^M R[\xi_\ell,\cc_\ell](\xx_i) 
      +\sum_{\xx_j \in \Gamma} (\nn(\xx_i) \otimes
      \nn(\xx_j))\eeta(\xx_j), \\
    \llambda_\ell &= \sum_{\xx_j \in \gamma_\ell} \eeta(\xx_j) 
      \Delta s_j, \\ 
    \mu_\ell &= \sum_{\xx_j \in \gamma_\ell}
      {(\xx_j - \cc_\ell)}^\perp \cdot \eeta(\xx_j) \Delta s_j,
  \end{aligned}
\end{equation}
for $i=1,\ldots,N$, $\ell=1,\ldots,M$.  We have applied the trapezoid
rule since all the integrands are smooth and periodic, and this
guarantees that the numerical solutions converges to the true solution
with spectral accuracy~\cite{tre-wei2014}.
